---
title: "Wk_4_hw_code"
---

### Read in Global Gender Inequality Data
### CSV to Spatial Data of the World
### Create New Column of Difference in Inequality Between 2010 & 2019

## Load Packages

```{r Packages}
library(tidyverse)
library(sf)
library(here)
library(janitor)
library(countrycode)
# country code changes country name to code --> for joining data later
library(plotly)
library(tmap)
library(tmaptools)
```

## Read in Data

Look in csv first --> see that na values = ..

Look in csv first --> can skip first 5 rows

```{r Data}
GII <- read_csv(here::here("Data", "Gender Inequality Index (GII).csv"),
                       locale = locale(encoding = "latin1"),
                       na = "..", skip=5)
World <- st_read(here::here("Data", "World_Countries__Generalized_.shp"))
```

## Data Wrangling
## Column Names

Select data we want(columns) and change country name to country code using (countrycode) package

We could also join it to country name, however joining to codes is preferable

```{r Columns}
GII_cols<- GII %>%
  clean_names()%>%
  select(country, x2019, x2010)%>%         # x2019, x2020 are the new column names(output from clean_names())
  mutate(difference=x2019-x2010)%>%        # make a new column (diff between 2019 and 2010 columns)
  slice(1:189,)%>%                         # look at data, see that from row 190 onward, there are aggregation data( areas, not countries) --> slice rows from 1 to 189. blank after comma means for all columns
  mutate(iso_code=countrycode(country, origin = 'country.name', destination = 'iso2c'))    # make new column, change name to country code. 'country' = column to look in and change to code
                                           # make new column --> call it iso_code
```

##Join

Joining csv to world shape file

```{r Join}
join_GII <- World %>% 
  clean_names() %>% 
  left_join(.,                             # . refers to World
            GII_cols,
            by = c("aff_iso" = "iso_code"))  
# aff_iso from World(shp)
# joining to
# iso_code from GII_cols (csv)
```

## Condition for Mapping

Composite measure reflecting inequality in achievement between women and men in 3 dimensions:
reproductive health, environment and the labour market
```{r Pre-Mapping}
# using case_when
# like an if statement
# mutate - make new column
# call it summary
# look at difference column, 
# when diff >0 (+ve number), write improved
# when diff ==0 (same), write same
# when diff is neither of those (TRUE), write worse
join_GII2 <- join_GII %>% 
  mutate(summary = case_when(difference > 0 ~ "improved",
                             difference == 0 ~ "the same",
                             TRUE ~ "worse"))

```

## Mapping

```{r Mapping}

tmap_mode("plot")
# qtm = quick thematic map
qtm(join_GII,
    fill="difference")
```

## MAPPING CONTINUED

More information on tm_polygons

https://rdrr.io/cran/tmap/man/tm_polygons.html

```{r Mapping Continued}

# defining own breaks
breaks=c(0.0,0.2,0.4,0.6,0.8,1.0)
diffbreaks=c(-0.4,-0.3,-0.2,-0.1,0,0.1)

# change GRCS to Mollewide
# preserves size and not direction, like WGS84
# direction is not important for this data
join_GII = st_transform (join_GII, crs = "+proj=moll")

# plot each map
# MAP a - 2019
tm1 <- tm_shape(join_GII)+
  tm_polygons("x2019",
              breaks = breaks,
              palette = "PuBu")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(a)", position = c(0, 0.85), size = 1.5)
# c = 0, 0.85
# 0 = x axis (so on the LHS)
# 0.85 = y axis ( on scale on 0 to 1- so near the top)

# MAP b - 2010
tm2 <- tm_shape(join_GII)+
  tm_polygons("x2010",
              breaks = breaks,
              palette = "PuBu")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(b)", position = c(0, 0.85), size = 1.5)

# Map c - diff
tm3 <- tm_shape(join_GII) + 
  tm_polygons("difference",
              #style="fixed",
              breaks=diffbreaks,
              palette=("Blues"),
               midpoint = NA) +  
  # Midpoint = The value mapped to the middle color of a diverging palette.
  # otherwise, will give error saying:
  # Warning: Breaks contains positive and negative values. Better is to use diverging scale instead, or set auto.palette.mapping to FALSE.
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(c)", position=c(0,0.85), size=1.5)

# legend
legend <- tm_shape(join_GII) +
    tm_polygons("difference",
                 palette=("Blues"),
               midpoint = NA) +
    #tm_scale_bar(position=c(0.2,0.04), text.size=0.6)+
    tm_compass(north=0, position=c(0.6,0.6))+
    tm_layout(legend.only = TRUE, legend.position=c(0.3,0.25),asp=0.1)+
    tm_credits("Mapped data:\nUN Gender Inequality Index\nWorld outline:\nArcGIS Hub 
               ", position=c(0.3,0.2), just="left")+
  tm_shape(join_GII) + 
  tm_polygons("x2019", 
              breaks=breaks,
              palette="PuBu")+
  tm_legend(show=TRUE)+
  tm_layout(legend.only = TRUE, legend.position=c(0.1,0.1),asp=0.1)

# in tmap can't make legends side by side unless use this workaround: 
# https://github.com/r-tmap/tmap/issues/476 

t=tmap_arrange(tm1, tm2, tm3, legend, ncol=2)
t

```

